Implications of afforestation policies for global markets

### 2023-05-31 paper.Rmd
Update paper.Rmd

# Introduction
Forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019]. With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation. In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019]. The gains that can be realized, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19]. At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland. Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

In addition, afforestation appears to not provide the same benefits in all regions of the world. In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015]. That contrasts with the results for tropical forests where forests there provide a cooling impact due to changes in both albedo and evapotranspiration [@li2015]. Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.

>>>>

# Introduction
Forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019]. With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation. In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019]. This is because of forests proven ability to sequester carbon not only the stems of trees [@davis2022] but their soils as well [@guo2021].  At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland. Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017]. Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.

The global gains that can be realized from afforestion, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman2019] to 205 gigatonnes [@bastin2019], as afforestation appears to not provide the same benefits in all regions of the world. In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015]. That contrasts with the results for tropical forests where forests there provide a cooling impact due to changes in both albedo and evapotranspiration [@li2015].
